SmartCheck: Automatic and Efficient Counterexample Reduction and Generalization\\$\;$\\DRAFT

### 2013-03-18 paper/paper.tex
paper: updated figure, wibbles.

In industrial uses, shrinking is essential.  Hughes, in describing commercial applications of QuickCheck, has noted that ``without it [shrinking], randomly generated
failing cases would often be so large as to be almost useless.''~\cite{qcjh}.
Hughes~\emph{et al.} give an extended example in which shrinking is essential in
debugging telecom software~\cite{telecom}.

@@

To motivate the benefits of SmartCheck, consider the graphs in
Figure~\ref{fig:graphs} and the Table~\ref{table:results}, giving the results
for using QuickCheck and SmartCheck on the program shown in
Figure~\ref{fig:initial} (we omit SmallCheck, since as noted, it is not even
feasible for the example).  We graph counterexample size and execution time.
Out of 2000 runs, we show the number of runs resulting in a shrunk
counterexample of a given size (by counting \ttp{N} constructors), and we show
the number of runs executing within a given time limit (in seconds).  Note that
we provide two execution-time graphs, since the scale of time scales differ
dramatically.  In the table, we show the mean, median, median, and the results
at the 95th percentile values.  (The curves produced by these values are not
necessarily Gaussian, so we omit standard the deviation.)

@@

We focus on \emph{algebraic data types}.  Algebraic data types are
represented as ``sums of products''.  We sometimes refer to the elements of a
sum type as a \emph{variant} that is \emph{tagged} by a constructor.  For
example, the type
%
\begin{code}
Maybe a = Just a | Nothing
\end{code}
%
\noindent
contains two variants, tagged by the constructors \ttp{Just} and \ttp{Nothing},
respectively.

@@

class Arbitrary a => SubTypes a where
  size     :: a -> Size
  index    :: a -> Idx -> Maybe SubVal
  replace  :: a -> Idx -> SubVal -> a
  constr   :: a -> String
  constrs  :: a -> [String]
  baseType :: a -> Bool
  subVals  :: a -> Tree SubVal
\end{code}
%
\noindent
The \ttp{SubTypes} type class requires QuickCheck's \ttp{Arbitrary} as a
super-class.  \ttp{SubTypes} has the following methods:
\begin{itemize}
\item \ttp{size} returns the size of a value---intuitively, the number of
  constructors contained within it,
\item \ttp{index} returns a \emph{sub-value} at an breadth-first index in a
  value,
\item \ttp{replace} replaces a sub-value at a particular focus (returning the
  original value if the index is out-of-bounds),
\item \ttp{constr} returns a string representation of the constructor tagging
  the value,
\item \ttp{constrs} returns the list of all possible constructors from the
  value's type,
\item \ttp{baseType} returns whether the type of the value is an ``interesting
  type'', which generally means it is user-defined and has positive arity.
  We discuss the purpose of base types in Section~\ref{sec:base}.
\item \ttp{subVals} returns a tree of all non base-type sub-values.  A tree is a
  value with the type
%
  \begin{code}
data Tree a = Node \{ rootLabel :: a
                   , subForest :: [Tree a] \}
  \end{code}
%
\end{itemize}

\noindent
To demonstrate typical evaluations of the methods, consider a binary tree type:
%
\begin{code}
data T = L | B T T
\end{code}
%
\noindent
and the value \ttp{tree}, labeled with indexes in a breadth-first order:
%
\begin{code}
tree = B\sub{0} (B\sub{1} L\sub{3}
             (B\sub{4} L\sub{6} L\sub{8}))
          (B\sub{2} L\sub{5} L\sub{7})
\end{code}
%
\noindent
Here are example applications of \ttp{SubTypes} methods; in the following, we
show the indexes with respect to the value \ttp{tree}:
%
\begin{code}
size tree = 9

@@

subVals (B (B L\sub{0} L\sub{1}) L\sub{2}) =
  Tree (B L\sub{0} L\sub{1}) [Tree L\sub{0} [], Tree L\sub{1} []]
       (Tree L\sub{1} [])
\end{code}
%
\noindent
The \ttp{SubVal} type is an existential data type, used as a generic container
for sub-values from a counterexample.  We will sometimes refer to the unwrapped
value returned by \ttp{index a i} as the \emph{$i$th sub-value of $a$}, so for
example, (\ttp{B\sub{4} L\sub{6} L\sub{8}}) is the 4th sub-value of \ttp{tree}.
An invariant of \ttp{index} is that for any value \ttp{a}, and for the smallest
$i \geq 0$ such that
%
\begin{code}
index a i == Nothing
\end{code}
%
\noindent
then for all $0 \leq j < i$,
%
\begin{code}
index a j /= Nothing
\end{code}
\noindent
We use this invariant as a termination case in recursive algorithms over the
sub-values of a value.

@@

Figure~\ref{fig:reduction} shows the reduction algorithm.  The function
\ttp{reduce} takes flags to customize the algorithm's behavior, a
counterexample \ttp{cex}, and the property \ttp{prop}.  At the beginning of
execution, we assume that a counterexample has been obtained somehow; our
implementation relies on QuickCheck as a ``front-end'' to generate a
counterexample that the algorithm automatically shrinks.  The reduction begins
at the first proper sub-value of \ttp{cex}; call it \ttp{v}.  When the index
\ttp{idx} becomes out-of-bounds and returns \ttp{Nothing}, the algorithm
terminates.  Otherwise, a list of new random values are generated.
%
\begin{code}
sizedArbitrary :: SubTypes a => a -> IO a
\end{code}
%
\noindent
generates a new value \ttp{v'} having the same type as \ttp{v} and that is
strictly smaller (with respect to the \ttp{size} method) than \ttp{v}.  Just
like QuickCheck's \ttp{arbitrary} method, \ttp{sizedArbitrary} generates
successfully larger counterexamples when generating new values with which to
replace a sub-value.

@@

For example, consider a simple calculator language containing constants,
addition, and division:
%
\begin{code}
data Exp = C Int
         | Add Exp Exp
         | Div Exp Exp
\end{code}
%
\noindent
Consider a program that takes an \ttp{Exp} as an argument that throws an error
if an evaluation of an \ttp{Exp} value contains a division-by-zero.  Testing for
the occurrence of the error, we might have a counterexample like the following:
%
\medskip%
\begin{sloppypar}
\small
\noindent%
\ttp{Add (Div (C 5) (C (-12))) (Add (Add (C 2) (C 4)) (Add (C 7) (Div (Add (C 7)
  (C 3)) (Add (C (-5)) (C 5)))))}
\end{sloppypar}
\medskip%
\noindent
In the counterexample, the culprit is a buried sub-value,
%
\begin{code}
Div (Add (C 7) (C 3)) (Add (C (-5)) (C 5))
\end{code}
%
Thus, when attempting to shrink an \ttp{Exp} value, it pays to test whether a
sub-value itself fails the property.

@@

\paragraph{Soundness}
The extrapolation algorithm is unsound in two ways.  First, it extrapolates from a set
of counterexamples to a universal claim, similar to QuickSpec~\cite{qs}.  By
tuning the parameters, the risk of an unsound generalization is reduced.
Second, note that each index is generalized with respect to the concrete
sub-values that exist in the counterexample.  So in some cases, a formula may be
returned that is overly general.  For example, consider the counterexample in
which both fields of \ttp{Add} contain values causing the failure:
%
\begin{code}
Add (Div (C 1) (C (-2)) (C 2))
    (Div (C 0) (C (-1)) (C 1))
\end{code}
%
\noindent
The universal generalization algorithm might return the formula
\noindent
%
\begin{code}
forall values x0 x1 . Add x0 x1
\end{code}
%
\noindent
The algorithm could be modified to so that when attempting to generalize a
sub-value at an index, all previously-generalized indexes are re-generalized
simultaneously, by taking the cross-product of random values for the current
index and for the previously-generalized indexes.  While this reduces the
potential for over-generalizing a counterexample, it is exponential in the
number of indexes generalized.  For universally quantified formulas, the reader
should read it as shorthand for quantifying each variable independently and
taking the conjunction of formulas.  For example, instead of
%
\begin{code}
forall values x0 x1 . Add x0 x1
\end{code}
%
\noindent
one should read
%
%
\begin{code}
      forall values x0 .
        Add x0 (Div (C 0) (C (-1)) (C 1))
  and forall values x1 .
        Add (Div (C 1) (C (-2)) (C 2)) x1
\end{code}

@@

We therefore introduce another kind of generalization we call \emph{existential
  sub-value generalization}.  In this generalization, if there a counterexample
containing every possible variant as a sub-value, then we abstract it.  For
example, the following are all counterexamples:
%
\begin{code}
Div (C 7) (Add (C (-5)) (C 5))
Div (Add (C 3) (C 2)) (C 0)
Div (C (-2)) (Div (C 0) (C 1))
\end{code}
%
\noindent
Because there are only three variants in the type, we have found counterexamples
built from each of them in the divisor.  We therefore can claim the following
formula holds:
%
\begin{code}
forall values x .
  forall constructors C .
    there exist arguments \(\stackrel{\rightarrow}{y}\) .
      such that Div x (C\(\stackrel{\rightarrow}{y}\))
\end{code}
%
\noindent
We therefore present an existential sub-value generalization algorithm that
performs constructor generalization.  Like with the other algorithms, this
algorithm also performs a breadth-first search over a counterexample.

@@

Figure~\ref{fig:cex-gen} illustrates the situation in which we have some values
and a property in the case that some values fail the property of the form
(\ttp{pre ==> post}).  Our goal is to find additional counterexamples in the
shaded region, in which \ttp{pre} is satisfied, \ttp{post} is not satisfied, and
the counterexample is not already characterized by a universal sub-value
generalization formula.  Iterating, we increasingly cover the shaded region
with new formulas until no more uncovered counterexamples exist or the testing
tool fails to find new counterexamples.

@@

Figure~\ref{fig:matches} show the shape-matching algorithm used for precondition
strengthening.  The function takes a candidate counterexample and a list of
previous counterexamples, together with the indexes at which they have been
generalized.  The basic idea of the algorithm is to determine whether two values
have the same ``shape''.  Intuitively, we consider two values to have the same
shape if in their respective tree representations, their constructors at each
node in the tree match, ignoring all base types (Section~\ref{sec:base}).
That is, intuitively, for values \ttp{a} and \ttp{b},
%
\begin{code}
subVals a == subVals b
\end{code}
%
\noindent
Furthermore, indexes that have been universally generalized (and all their
children) match any value, since a universally generalized index indicates that
counterexamples have been found for any possible constructor.  In our
implementation, we also consider existentially generalized indexes to match any
value.  Doing so is a design choice that is more aggressive about covering the
space of counterexamples at the risk of omitting some.

@@

SmartCheck does not implement a counterexample discovery algorithm.  While any
method can be used to discover an initial counterexample, SmartCheck by default
uses QuickCheck to discover an initial counterexample.

@@

SmartCheck returns a smaller counterexample, and then generalizes it:
%
\begin{code}
forall values x0 x1 x2 x3:
  StackSet
    (Screen (Workspace x0 (-1) (Just x1)) 1 1)
    x2 x3 (fromList [])
\end{code}
%
\noindent
We can look at what was generalized and what was not.  First, the last element
of \ttp{StackSet} is a map, and we treated it as a base type, so it is not
analyzed at all.  But it is an empty map.  The second and third fields of
\ttp{StackSet} are generalized, but the first one is not.  So there is something
particular about it.  So the culprit is one of the small constants (1 and -1) or
having a \ttp{Just} value rather than a \ttp{Nothing}: it turns out that what
matters is having a \ttp{Just} value, which is the stack field that deletion
works on.

@@

\subsection{Natural Language Processing}
In 2012, a question was posted on the programming message board
Stack Overflow asking about how to shrinking large nested data
types.\footnote{\url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}}
The poster writes:
\begin{quote}
$\ldots$ I tend to get an incomprehensible page full of output. $\ldots$
  Implementing the shrink function for each of my types seems to help a little,
  but not as much as I'd like.  $\ldots$ If I try to tune my shrink
  implementations, I also find that QC starts taking a very long time.
\end{quote}
\noindent
The question relates to the Geni natural language processing (NLP) package
implemented in Haskell~\cite{kow}.  Specifically, counterexamples to a property
attempting to show that a macro expansion function is its own inverse are
enormous, requiring 200-300 80-character lines to print.

@@

Counterexample reduction in general is a relatively unexplored topic, but there
has been some treatment of the topic.  Zeller and Hildebrandt describe an
application of greedy search to shrink counterexamples they call
``delta-debugging'' (DD)~\cite{dd}.  The authors apply their work shrinking HTML
inputs to crash Mozilla and shrinking C programs to trigger a bug in GCC.
Subsequent generalizations are reported by Misherghi and Su in which they
perform greedy search on tree-structured data; they call their approach
hierarchical delta-debugging (HDD)~\cite{hdd}.

@@

SmartCheck bears some similarity to QuickSpec, a testing-based library that
infers equational properties about programs~\cite{qs} insofar as they both
attempt to generalize counterexamples based on specific inputs.

@@

Another aspect of displaying large counterexamples that we have not explored is
to exploit sharing.  Constructs might be repeated that can be abstracted out.
For example, instead of a counterexample like
%
\begin{code}
Add (Div (C 1) (C (-2)) (C 2))
    (Div (C 1) (C (-1)) (C 1))
\end{code}
%
\noindent
we might instead return
%
\begin{code}
Add (div (-2) 2) (div (-1) 1)
  where div x y = Div (C 1) (C x) (C y)
\end{code}
%
\noindent
Discovering and exploiting sharing is future work.

>>>>

In industrial uses, shrinking is essential.  In describing commercial
applications of QuickCheck, Hughes has noted that ``without it [shrinking],
randomly generated failing cases would often be so large as to be almost
useless.''~\cite{qcjh}.  Hughes~\emph{et al.} also give an extended example in
which shrinking is essential in debugging telecom software~\cite{telecom}.

@@

To motivate the benefits of SmartCheck, consider the graphs in
Figure~\ref{fig:graphs} and the Table~\ref{table:results}, giving the results
for using QuickCheck and SmartCheck on the program shown in
Figure~\ref{fig:initial} (we omit SmallCheck, since as noted, it is not even
feasible for the example).  We graph counterexample size and execution time.
Out of 2000 runs, we show the number of runs resulting in a shrunk
counterexample of a given size (by counting \ttp{N} constructors), and we show
the number of runs executing within a given time limit (in seconds).  Note that
we provide two execution-time graphs, since the scale of time scales differ
dramatically.  In the table, we show the mean, median, median, and the results
at the 95th percentile values.  (The distributions produced by these values are
not necessarily Gaussian, so we omit standard the deviation.)

@@

We focus on algebraic data types represented as ``sums of products''.  We
sometimes refer to the elements of a sum type as a \emph{variant} that is
\emph{tagged} by a constructor.  For example, the type
%
\begin{code}
Maybe a = Just a | Nothing
\end{code}
%
\noindent
contains two variants, tagged by the constructors \ttp{Just} and \ttp{Nothing},
respectively.

@@

class Arbitrary a => SubTypes a where
  size     :: a -> Size
  index    :: a -> Idx -> Maybe SubVal
  replace  :: a -> Idx -> SubVal -> a
  constr   :: a -> String
  constrs  :: a -> [String]
  baseType :: a -> Bool
  subVals  :: a -> Tree SubVal
\end{code}
%
\noindent
The \ttp{SubTypes} type class requires QuickCheck's \ttp{Arbitrary} as a
super-class.  \ttp{SubTypes} has the following methods:
\begin{itemize}
\item \ttp{size} returns the size of a value---intuitively, the number of
  constructors contained within it,
\item \ttp{index} returns a \emph{sub-value} at an breadth-first index in a
  value,
\item \ttp{replace} replaces a sub-value at a particular focus (returning the
  original value if the index is out-of-bounds),
\item \ttp{constr} returns a string representation of the constructor tagging
  the value,
\item \ttp{constrs} returns the list of all possible constructors from the
  value's type,
\item \ttp{baseType} is true when the type of the value is an ``uninteresting
  type'', which generally means it is user-defined and has positive arity.
  We discuss the purpose of base types in Section~\ref{sec:base}.
\item \ttp{subVals} returns a tree of all non base-type sub-values.  A tree is a
  value with the type
%
  \begin{code}
data Tree a = Node \{ rootLabel :: a
                   , subForest :: [Tree a] \}
  \end{code}
%
\end{itemize}

\noindent
To illustrate typical evaluations of the methods, consider a binary tree type:
%
\begin{code}
data T = L | B T T
\end{code}
%
\noindent
and the value \ttp{tree}, labeled with indexes in a breadth-first order:
%
\begin{code}
tree = B\sub{0} (B\sub{1} L\sub{3}
             (B\sub{4} L\sub{6} L\sub{8}))
          (B\sub{2} L\sub{5} L\sub{7})
\end{code}
%
\noindent
Here are example applications of \ttp{SubTypes} methods; in the following, we
show the indexes with respect to the value \ttp{tree}:
%
\begin{code}
size tree = 9

@@

subVals (B (B L\sub{0} L\sub{1}) L\sub{2}) =
  Node (B L\sub{0} L\sub{1}) [Node L\sub{0} [], Node L\sub{1} []]
       (Node L\sub{1} [])
\end{code}
%
\noindent
The \ttp{SubVal} type is an existential data type, used as a generic container
for sub-values from a counterexample.  We will sometimes refer to the unwrapped
value returned by \ttp{index a i} as the \emph{$i$th sub-value of $a$}, so for
example, (\ttp{B\sub{4} L\sub{6} L\sub{8}}) is the 4th sub-value of \ttp{tree}.
An invariant of \ttp{index} is that for any value \ttp{a}, and for the smallest
$i \geq 0$ such that
%
\begin{code}
index a i == Nothing
\end{code}
%
\noindent
then for all $0 \leq j < i$,
%
\begin{code}
index a j /= Nothing
\end{code}
\noindent
We use this invariant as a termination case in recursive algorithms over the
sub-values of a value.

@@

Figure~\ref{fig:reduction} shows the reduction algorithm.  The function
\ttp{reduce} takes flags to customize the algorithm's behavior, a counterexample
\ttp{cex}, and the property \ttp{prop}.  At the beginning of execution, we
assume that a counterexample has been obtained somehow; our implementation
relies on QuickCheck to generate an initial counterexample and to generate
arbitrary well-typed random values (by the \ttp{arbitrary} method).  The
reduction begins at the first proper sub-value of \ttp{cex}; call it \ttp{v}.
When the index \ttp{idx} becomes out-of-bounds and returns \ttp{Nothing}, the
algorithm terminates.  Otherwise, a list of new random values are generated.
%
\begin{code}
sizedArbitrary :: SubTypes a => a -> IO a
\end{code}
%
\noindent
generates a new value \ttp{v'} having the same type as \ttp{v} and that is
strictly smaller (with respect to the \ttp{size} method) than \ttp{v}.  Just
like QuickCheck's \ttp{arbitrary} method, \ttp{sizedArbitrary} generates
successively larger counterexamples when generating new values with which to
replace a sub-value.

@@

For example, consider a simple calculator language containing constants,
addition, and division, together with an evaluator that checks if the divisor is
0 and returning \ttp{Nothing} in that case:
%
\begin{code}
data Exp = C Int
         | Add Exp Exp
         | Div Exp Exp

eval :: Exp -> Maybe Int
eval (C i) = Just i
eval (Add e0 e1) =
  liftM2 (+) (eval e0) (eval e1)
eval (Div e0 e1) =
  let e = eval e1 in
  if e == Just 0 then Nothing
    else liftM2 div (eval e0) e
\end{code}
%
\noindent
Now consider the property \ttp{prop\_div}, claiming that if \ttp{divSubTerms}
holds on an expression, then the evaluator returns \ttp{Just} a value:
%
\begin{code}
divSubTerms :: Exp -> Bool
divSubTerms (C _)         = True
divSubTerms (Div _ (C 0)) = False
divSubTerms (Add e0 e1)   =  divSubTerms e0
                          && divSubTerms e1
divSubTerms (Div e0 e1)   = divSubTerms e0
                          && divSubTerms e1

prop_div e = divSubTerms e ==> eval e /= Nothing
\end{code}
%
Testing \ttp{prop\_div}, we might have a counterexample like the following:
%
\medskip%
\begin{sloppypar}
\small
\noindent%
\ttp{Add (Div (C 5) (C (-12))) (Add (Add (C 2) (C 4)) (Add (C 7) (Div (Add (C 7)
  (C 3)) (Add (C (-5)) (C 5)))))}
\end{sloppypar}
\medskip%
\noindent
In the counterexample, the culprit is a buried sub-value,
%
\begin{code}
Div (Add (C 7) (C 3)) (Add (C (-5)) (C 5))
\end{code}
%
Thus, when attempting to shrink an \ttp{Exp} value, it pays to test whether a
sub-value itself fails the property.

@@

\paragraph{Soundness}
The extrapolation algorithm is unsound in two ways.  First, it extrapolates from
a set of counterexamples to a universal claim, similar to QuickSpec~\cite{qs}.
By tuning the parameters, the risk of an unsound generalization is reduced by
requiring more or larger values to fail the property.

Second, note that each index is generalized with respect to the concrete
sub-values that exist in the counterexample.  So in some cases, a formula may be
returned that is overly general.  For example, consider the counterexample in
which both fields of \ttp{Add} contain values causing the failure:
%
\begin{code}
Add (Div (C 1) (C (-2)) (C 2))
    (Div (C 0) (C (-1)) (C 1))
\end{code}
%
\noindent
since no matter what random value the first field of \ttp{Add} is replaced with,
the property fails by the second field and vice versa for the second.  The
universal generalization algorithm might return the formula
\noindent
%
\begin{code}
forall values x0 x1 . Add x0 x1
\end{code}
%
\noindent
The algorithm could be modified so that when attempting to generalize a
sub-value at an index, all previously-generalized indexes are re-generalized
simultaneously, by taking the cross-product of random values for the current
index and for the previously-generalized indexes.  While this reduces the
potential for over-generalizing a counterexample, it is exponential in the
number of indexes generalized.  For universally quantified formulas, the reader
should read it as shorthand for quantifying each variable independently and
taking the conjunction of formulas.  For example, instead of
%
\begin{code}
forall values x0 x1 . Add x0 x1
\end{code}
%
\noindent
one should read
%
%
\begin{code}
      forall values x0 .
        Add x0 (Div (C 0) (C (-1)) (C 1))
  and forall values x1 .
        Add (Div (C 1) (C (-2)) (C 2)) x1
\end{code}

@@

We therefore introduce another kind of generalization we call \emph{existential
  sub-value generalization}.  In this generalization, if there a counterexample
containing every possible variant as a sub-value, then we abstract it.  For
example, the following are all counterexamples:
%
\begin{code}
Div (C 7) (Add (C (-5)) (C 5))
Div (Add (C 3) (C 2)) (C 0)
Div (C (-2)) (Div (C 0) (C 1))
\end{code}
%
\noindent
Because there are only three variants in the type, we have found counterexamples
built from each of them in the divisor.  We therefore can claim the following
formula holds:
%
\begin{code}
forall values x .
  forall constructors c .
    there exist arguments \(\stackrel{\rightarrow}{y}\) .
      such that Div x (c\(\stackrel{\rightarrow}{y}\))
\end{code}
%
\noindent
We therefore present an existential sub-value generalization algorithm that
performs constructor generalization.  Like with the other algorithms, this
algorithm also performs a breadth-first search over a counterexample.

@@

Figure~\ref{fig:cex-gen} illustrates a property of the form (\ttp{pre ==>
  post}).  Points are specific counterexamples

 the situation in which we have some values
and a property in the case that some values fail the property of the form
.  Our goal is to find additional counterexamples in the
shaded region, in which \ttp{pre} is satisfied, \ttp{post} is not satisfied, and
the counterexample is not already characterized by a universal sub-value
generalization formula.  Iterating, we increasingly cover the shaded region
with new formulas until no more uncovered counterexamples exist or the testing
tool fails to find new counterexamples.

@@

Figure~\ref{fig:matches} shows the shape-matching algorithm used for precondition
strengthening.  The function takes a candidate counterexample and a list of
previous counterexamples, together with the indexes at which they have been
generalized.  The basic idea of the algorithm is to determine whether two values
have the same ``shape''.  Intuitively, we consider two values to have the same
shape if in their respective tree representations, their constructors at each
node in the tree match, ignoring all base types (Section~\ref{sec:base}).
That is, intuitively, for values \ttp{a} and \ttp{b},
%
\begin{code}
subVals a == subVals b
\end{code}
%
\noindent
Furthermore, indexes that have been universally generalized (and all their
children) match any value, since a universally generalized index indicates that
counterexamples have been found for any possible constructor.  In our
implementation, we also consider existentially generalized indexes to match any
value.  Doing so is a design choice that is more aggressive about covering the
space of counterexamples at the risk of omitting some.

@@

SmartCheck does not implement a counterexample discovery algorithm or an
\ttp{arbitrary} method.  Any initial counterexample can be passed in or
generated, and the algorithm depends on \ttp{arbitrary} to generate new
well-typed values.

@@

SmartCheck returns a smaller counterexample, and then generalizes it:
%
\begin{code}
forall values x0 x1 x2 x3:
  StackSet
    (Screen (Workspace x0 (-1) (Just x1)) 1 1)
    x2 x3 (fromList [])
\end{code}
%
\noindent
Let us examine what was generalized and what was not.  First, the last element
of \ttp{StackSet} is a map, and we treated it as a base type, so it is not
analyzed at all.  But it is an empty map.  The second and third fields of
\ttp{StackSet} are generalized, but the first one is not.  So there is something
particular about it.  So the culprit is one of the small constants (1 and -1) or
having a \ttp{Just} value rather than a \ttp{Nothing}: it turns out that what
matters is having a \ttp{Just} value, which is the stack field that deletion
works on.

@@

\subsection{Natural Language Processing}
In 2012, a question was posted on the programming message board
Stack Overflow asking about how to shrink large nested data
types.\footnote{\url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}}
The poster writes:
\begin{quote}
$\ldots$ I tend to get an incomprehensible page full of output. $\ldots$
  Implementing the shrink function for each of my types seems to help a little,
  but not as much as I'd like.  $\ldots$ If I try to tune my shrink
  implementations, I also find that QC starts taking a very long time.
\end{quote}
\noindent
The question relates to the Geni natural language processing (NLP) package
implemented in Haskell~\cite{kow}.  Specifically, counterexamples to a property
attempting to show that a macro expansion function is its own inverse are
enormous, requiring 200-300 80-character lines to print.

@@

A few other researchers have also investigated counterexample shrinking.  Zeller
and Hildebrandt describe an application of greedy search to shrink
counterexamples they call ``delta-debugging'' (DD)~\cite{dd}.  The authors apply
their work shrinking HTML inputs to crash Mozilla and shrinking C programs to
trigger a bug in GCC.  Subsequent generalizations are reported by Misherghi and
Su in which they perform greedy search on tree-structured data; they call their
approach hierarchical delta-debugging (HDD)~\cite{hdd}.

@@

Finally, SmartCheck bears some similarity to QuickSpec, a testing-based library
that infers equational properties about programs~\cite{qs} insofar as they both
attempt to generalize counterexamples based on specific inputs.

@@

Another aspect of displaying large counterexamples that we have not explored is
to exploit sharing.  Constructs might be repeated that can be abstracted out.
For example, instead of a counterexample like
%
\begin{code}
Add (Div (C 1) (C (-2)) (C 2))
    (Div (C 1) (C (-1)) (C 1))
\end{code}
%
\noindent
we might instead return
%
\begin{code}
Add (div (-2) 2) (div (-1) 1)
  where div x y = Div (C x) (C y)
\end{code}
%
\noindent
Discovering and exploiting sharing is future work.
